### 2011-12-25 trunk/GUIErrorDetector/paper/technique.tex
paper tweak

\section{Technique}

We first gives our high-level approach formulation, then present
an error detection algorithm. Finally, we show how to instantiate
the error detection to different GUI frameworks.

@@

\noindent {\textsc{\textbf{Definition 1 (UI Thread).}}} {UI thread
is a special thread created by the GUI framework during
GUI initialization. After the GUI becomes visible, the UI thread
takes charge of the application, handle events from the GUIs,
or spawn new threads to process lengthy operations in the background. }\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 1.}}} {We assume that each multithreaded
GUI application has a single global UI thread. This is true for applications
built on top of GUI frameworks that adapt the single-thread rule. It is
possible that an application launches another application with its own UI thread
by initializing a different process. In that case, we require the new
launched application to be analyzed separately.}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 2 (non-UI Thread).}}} {Any other
threads except for the UI thread is called non-UI threads.}\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 2.}}} { We assume that each non-UI
thread can only be spawned by the UI thread. Under this assumption, all post-initialization
GUI work naturally occurs in the UI-thread. Once the GUI is visible, the
application is driven by events, which are always handled in the UI thread. We
believe this assumption is reasonable, since if a non-UI
thread is spawned during pre-initialization GUI work and that thread accesses
a GUI object, an exception becomes apparent and the whole application may abort. This
is unlikely for most of the realistic GUI applications.
}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 3 (UI Accessing Methods).}}} { A method
whose execution may read or write a UI object is called a UI Accessing Method.}\vspace{1mm}

@@

\noindent  {\textsc{\textbf{Definition 5 (Invalid Thread Access).}}} {An invalid
thread access error occurs if the UI thread spawns a non-UI thread, and there
exists a path from the non-UI thread's \CodeIn{start} method to any UI accessing methods
without going through any safe methods. }\vspace{2mm}

@@

We cast the problem of finding invalid thread

Choose a right abstraction, why call graph is good?

a call graph is a good approximation

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that coomputing a sound
call graph for partial programs such as an eclipse plugin,
or in the presence of reflection, is non-trivial. Existing
call graph construction algorithms either suffer from incompleteness
or imprecision.  XX We describe our combined xxx in Section XXX.

We also note that static analysis may report redundant

\begin{figure}[t]
\textbf{Input}: a Java program $P$\\
\textbf{Output}: a set of potential $errors$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\STATE errors $\leftarrow$ $\emptyset$
\STATE cg $\leftarrow$ $constructCallGraph$($P$)
\STATE entryNodes $\leftarrow$ $getEntryNodes$(cg)
\STATE uiAccessNodes $\leftarrow$ $getUIAccessNodes$(cg)
\STATE safeNodes $\leftarrow$ $getSafeNodes$(cg)
\FOR{each entryNode in entryNodes}
\STATE threadStartNodes $\leftarrow$ $getReachableStartsByBFS$(cg)
\FOR{each threadStart in threadStartNodes}
\STATE queue $\leftarrow$ an empty queue
\STATE queue.enqueueAll(threadStart.$getAllSuccNodes$())
\WHILE{queue.isNotEmpty()}
\STATE nextNode $\leftarrow$ queue.dequeue()
\FOR{each succNode in cg.$getSuccNodes$(nextNode)}
\IF{succNode $\in$ uiAccessNodes}
\STATE errors $\leftarrow$ errors $\cup$ $createErrorReport$(succNode)
\ELSIF{succNode $\in$ safeNodes}
\STATE continue
\ELSE
\STATE queue.enqueue(succNode)
\ENDIF 
\ENDFOR
\ENDWHILE
\ENDFOR
\ENDFOR
\RETURN $errors$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{Algorithm for detecting errors in multithreaded GUI programs. 
Four utility methods in lines 2 -- 4 are customized for each GUI platforms
(not each program) as described in Section~\ref{sec:platforms}.
} \label{fig:detectalgorithm}
\end{figure}

\subsubsection{Call Graph Construction}

@@

1. public class MyActivity extends Activity \{
2.    final List<View> cachedViews = new LinkedList<View>();
3.    @Override
4.    public void onCreate(Bundle savedInstanceState) \{
5.        super.onCreate(savedInstanceState);
6.        setContentView(R.layout.main);
7.        Button button = (Button) findViewById(R.id.\textbf{button\_id});
8.        button.setOnClickListener(new Button.OnClickListener() \{
9.            @Override
10.           public void onClick(View v) \{
11.               button.setText("Button Clicked.");
12.               cachedViews.add(v);
13.           \}
14.       \});
15.   \}
16. \}
\end{alltt}
\end{CodeOut}
\label{fig:sampleandroid}
\caption{A Sample Android Code ..}
\end{figure}

@@

\begin{figure}[t]
\textbf{Input}: a Java program $P$, a list of classes $cls$\\
\textbf{Output}: a call graph $cg$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\STATE callSiteMap $\leftarrow$ new Map$\langle$CallSite, List$\langle$Type$\rangle$$\rangle$
\STATE rta\_cg $\leftarrow$ $constructCallGraphByRTA$($P$)
\FOR{each $method$ in $P$}
\FOR{each  $callSite$ in $method$}
\STATE type $\leftarrow$ $staticType$(callSite)
\IF{type $\in$ $cls$}
\FOR{each subType of $subType$(type)}
\IF{$hasObjectCreated$(subType, rta\_cg)}
\STATE callSiteMap[callSite].add(subType)
\ENDIF
\ENDFOR
\ENDIF
\ENDFOR
\ENDFOR
\STATE $cg$ $\leftarrow$ $constructCallGraphByKCFA$($P$, callSiteMap)
\RETURN $cg$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{Algorithm for constructing a call graph using a combination of the RTA~\cite{rta} and k-CFA~\cite{kcfa} algorithms.
$cls$ is a list of class XXXX, whose invocatioon is treated using RTA; while the remaining call graph
part is constructed using k-CFA.
} \label{fig:cgalgorithm}
\end{figure}

\subsubsection{Heuristic Filters and Annotations}
(XXX)

>>>>

\section{Technique}
\label{sec:technique}

We first gives our high-level approach formulation, then present
an error detection algorithm. Finally, we show how to instantiate
the error algorithm to different GUI frameworks.

@@

\noindent {\textsc{\textbf{Definition 1 (UI Thread).}}} {UI thread
is a special thread created by the GUI framework during
GUI initialization. After the GUI becomes visible, the UI thread
takes charge of the application to handle events from the GUIs,
or spawn new threads to process lengthy operations in the background. }\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 1.}}} {We assume that each multithreaded
GUI application has a single global UI thread. This is true for applications
built on top of GUI frameworks adapting the single-thread rule. The
only exception is that 
 an application may folk a new process to launch another
application with its own UI thread. In that case, we require the new
launched application to be analyzed separately.}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 2 (non-UI Thread).}}} {Any other
threads except for the UI thread in a multithreaded GUI application
 are called non-UI threads.}\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 2.}}} { We assume that each non-UI
thread can only be spawned by the UI thread. Under this assumption,
 all post-initialization
GUI work naturally occurs in the UI-thread. Once the GUI is visible, the
application is driven by events, which are always handled in the UI thread.
We can safely ignore non-UI threads that are spawned during pre-initilaizaiton
GUI work, since if such threads access a GUI object, an exception becomes
immediately apparent and the whole application may abort even before the
GUI is visible. This is highly unlikely for most of the realistic GUI
applications.
}\vspace{1mm}

\noindent {\textsc{\textbf{Definition 3 (UI Accessing Methods).}}} { A method
whose execution may read or write a UI object is called a UI accessing method.}\vspace{1mm}

@@

\noindent  {\textsc{\textbf{Definition 5 (Invalid Thread Access).}}} {An invalid
thread access error occurs when the UI thread spawns a non-UI thread, and there
exists a path from the non-UI thread's \CodeIn{start} method to any UI accessing methods
without going through any safe methods. }\vspace{2mm}

@@

The algorithm for detecting potential invalid thread access errors
is shown in Figure~\ref{fig:detectalgorithm}. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specially, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretical ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative static call graph is a superset of the ideal call graph;
it over-approximates the
dynamic call graph of every possible execution. 
We believe call graph has a suitable granularity for this problem while
Compared with other
program representations like control flow graph or program dependence
graph. That is because, in an object-oriented programming language like
Java, a method is the basic unit of a program. Thus, reporting errors
at the method level can be sufficient in most cases.

Our algorithm in Figure~\ref{fig:detectalgorithm} works as follows. It
constructs a static call graph for the tested program (line 2),
and specifies entry nodes for it. The algorithm starts to perform
call graph traverse from every entry node (line 6). For every entry node,
it first uses Breadth First Search (BFS) to find all reachable
\CodeIn{Thread.start()} method (line 7). Each \CodeIn{Thread.start()} method
indicates a new, non-UI thread is spawned.  Then, from each
\CodeIn{Thread.start()} node, the algorithm uses a slightly modified 
BFS algorithm to find reachable UI accessing methods (lines 9 -- 23).
Specifially, it uses a queue to keep track of all visited nodes, and
the algorithm checks every successive node to be visited;
it reports an error if that node
is a UI accessing method (line 15), or skips it if it is
a safe method (line 17), or simply keeps traversing (line 19).

If the algorithm determines that a UI accessing method may be invoked
by a non-UI thread, it creates an error report containing the full
method call chain from the entry node (line 15, the \CodeIn{createErrorReport}
method). An example call chain for
the motivating example in Figure~\ref{fig:androiderror} is shown
in Figure~\ref{fig:report}.

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph for partial programs such as an eclipse plugin,
or in the presence of reflection, is non-trivial. Existing
call graph construction algorithms either suffer from incompleteness
or imprecision. GUI applications built on some popular frameworks
often use configuration files and reflection to specify GUI layout
or extension points, therefore, reflection must be properly handled.
We developed a combined RTA
and k-CFA call graph construction algorithm in Section~\ref{sec:cg}
to alleviate this problem.

We also note that static analysis may report false positives due
to its conservative nature, and multiple warnings that acutally corrspond to
the same error. In addition, some GUI applications may even use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
must be considered to make the call graph sound. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and remove redundant reports, and annotation
support for programmers to specify call relationship for native methods.

\begin{figure}[t]
\textbf{Input}: a Java program $P$\\
\textbf{Output}: a set of potential $errors$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\STATE errors $\leftarrow$ $\emptyset$
\STATE cg $\leftarrow$ $constructCallGraph$($P$)
\STATE entryNodes $\leftarrow$ $getEntryNodes$(cg)
\STATE uiAccessNodes $\leftarrow$ $getUIAccessNodes$(cg)
\STATE safeNodes $\leftarrow$ $getSafeNodes$(cg)
\FOR{each entryNode in entryNodes}
\STATE threadStartNodes $\leftarrow$ $getReachableStartsByBFS$(cg)
\FOR{each threadStart in threadStartNodes}
\STATE queue $\leftarrow$ an empty queue
\STATE queue.enqueueAll(threadStart.$getAllSuccNodes$())
\WHILE{queue.isNotEmpty()}
\STATE nextNode $\leftarrow$ queue.dequeue()
\FOR{each succNode in cg.$getSuccNodes$(nextNode)}
\IF{succNode $\in$ uiAccessNodes}
\STATE errors $\leftarrow$ errors $\cup$ $createErrorReport$(succNode)
\ELSIF{succNode $\in$ safeNodes}
\STATE continue
\ELSE
\STATE queue.enqueue(succNode)
\ENDIF 
\ENDFOR
\ENDWHILE
\ENDFOR
\ENDFOR
\RETURN $errors$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{Algorithm for detecting invalid thread access errors in multithreaded GUI programs. 
Any call graph construction algorithm can be used (line 2).
Three utility methods in lines 3 -- 5 are specific for each GUI framework
 as described in Section~\ref{sec:platforms}.
} \label{fig:detectalgorithm}
\end{figure}

\subsubsection{Call Graph Construction}
\label{sec:cg}

1. public class MyActivity extends Activity \{
2.    final List<View> cachedViews = new LinkedList<View>();
3.    @Override
4.    public void onCreate(Bundle savedInstanceState) \{
5.        super.onCreate(savedInstanceState);
6.        setContentView(R.layout.main);
7.        Button button = (Button) findViewById(R.id.\textbf{button\_id});
8.        button.setOnClickListener(new Button.OnClickListener() \{
9.            @Override
10.           public void onClick(View v) \{
11.               button.setText("Button Clicked.");
12.               cachedViews.add(v);
13.           \}
14.       \});
15.   \}
16. \}
\end{alltt}
\end{CodeOut}
\label{fig:sampleandroid}
\caption{Sample GUI application code on the Android platform. The layout
XML file (the top) specifies a Button object declaratively, 
and the Java code (the bottom)
first loads the XML file (line 6) and then uses reflection to create
a Button object by its ID (line 7, the \CodeIn{findViewByd} method).}
\end{figure}

In the context of object-oriened languages wth dynamic dispatch, the
crucial step in constructing a call graph is to compute a conservative
approximation of the set of methods that can be invoked by a given
virtual method call. Call graph construction algorithms have been
studied intensively in the 1990s. While their original formulations
use a variety of formalisms, most of them can be recast as set-based
analyses. Various algorithms differ primarily in the number of sets
that are used to approximate run-time values of expressions. 
In practice, algorithms such as RTA that use a single set for the whole
program scale well but achieves relatively lower precision. The
k-CFA algorithms maintains one set per expression achieves better
precision than RTA.

Unfortunately, for GUI applications that use reflection, we found both
RTA and k-CFA is insufficient to build a satisfactory call graph. Here
is an example in Figure~\ref{fig:sampleandroid} to illustrate their limitations.

(XXX)

@@

@@

\begin{figure}[t]
\textbf{Input}: a Java program $P$, a list of classes $cls$\\
\textbf{Output}: a call graph $cg$\\
\vspace{-5mm}
\begin{algorithmic}[1]
\STATE callSiteMap $\leftarrow$ new Map$\langle$CallSite, List$\langle$Type$\rangle$$\rangle$
\STATE rta\_cg $\leftarrow$ $constructCallGraphByRTA$($P$)
\FOR{each $method$ in $P$}
\FOR{each  $callSite$ in $method$}
\STATE type $\leftarrow$ $staticType$(callSite)
\IF{type $\in$ $cls$}
\FOR{each subType of $subType$(type)}
\IF{$hasObjectCreated$(subType, rta\_cg)}
\STATE callSiteMap[callSite].add(subType)
\ENDIF
\ENDFOR
\ENDIF
\ENDFOR
\ENDFOR
\STATE $cg$ $\leftarrow$ $constructCallGraphByKCFA$($P$, callSiteMap)
\RETURN $cg$
%\ENDWHILE
\vspace{-2mm}
\end{algorithmic}
\caption{A combined  RTA~\cite{rta} and k-CFA~\cite{kcfa} call graph construction algorithm.
$cls$ is a list of classes whose method invocations are added to the
call graph using the RTA algorithm; while method invocations from other classes
are added to the call graph using the k-CFA algorithm.
} \label{fig:cgalgorithm}
\end{figure}

\subsubsection{Heuristic Filters and Annotations}
\label{sec:heuristic}
